---
title: "Compare 3 Clustering methods for Method Consensus"
author: "Kelly Brewer, HC clustering by MD and RS"
date: "11/1/21"
output: 
  html_document:
    toc: true
    toc_depth: 3
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file uses 3 clustering methods to compare which cluster group each of the PCOS cohort subjects would fall into given each of the 3 methods. Cluster methods use 8 traits and include:
  1) Hierarchical Clustering (stats package)- connectivity-based clustering
  2) K-means Clustering (flexclust package)- centroid-based clustering
  3) Gaussian mixture models (ClusterR package)- distribution-based clustering

***
## Input Data Fields and Structure
```{r read_data, eval=TRUE, echo=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
library(ggrepel)
library(tidymodels)
library(reshape2)
library(factoextra)
library(FactoMineR)
library(ClusterR)
library(fpc)
library(stats)
library(gplots)
library(clusterSim)
library(describedata)
library(lattice)
library(corrplot)
library(DT)
library(ggradar)
library(qwraps2)
options(qwraps2_markup = "markdown")
library(Hmisc)
library(ggstatsplot)


################################################
# Functions                                    #
################################################
## Reverse normal transformation
rntransform <- function(y) {
  out <- rank(y)
  out[is.na(y)] <- NA
  qnorm((out - 0.5) / max(out, na.rm=T))
}

################################################
# Variables                                    #
################################################

## Path/file of the input file, which must have columns according to
## cluster_input_cols below.
input_file <- "path/to/SampleData_355.txt"

## Names of sample ID, age, bmi, and the other 7 traits
sample <- 'sample_id'; age <- 'age'; bmi <- 'bmi'
traitnames = c('T', 'dheas', 'i0', 'g0', 'shbg', 'lh', 'fsh')

## variable names for displaying/plotting results
var_labels <- c('BMI', 'T', 'DHEAS', 'Ins0', 'Glu0', 'SHBG', 'LH', 'FSH')

## The assay method for each of the 7 traits is another variable, named by
## adding a postfix to the trait name.  The postfix in our data is
## '_assay_method'.  For example, dheas_assay_method.
assay <- '_assay_method'

## columns of the input file
cluster_input_cols <- c(sample, age, bmi,
                        traitnames,
                        paste0(traitnames, assay))
## columns of variables (some to be defined later) for clustering
d <- '.'
cluster_cols <- c(sample, age,
                  paste(c(bmi, traitnames), 'z', sep=d))

## Methods for distance calculation and for hierarchical clustering.
## Our initial analysis used manhattan distance and ward.D clustering method.
## ward.D2 is probably a better choice.
dist_metric <- 'manhattan'
clust_method <- 'ward.D'

## Read in input file (must have columns according to cluster_input_cols)
m.df0 <- read.delim(input_file, na.strings=c("", "#N/A", "missing"))
str(m.df0)

```
***

## Remove duplicates, outliers and missing data
```{r remove_dupes_incomplete, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}
## Remove duplicate entries (keep the first occurrence for each set of duplicates)
dupidx = duplicated(m.df0[, cluster_input_cols[1:(3+length(traitnames))]])
if(sum(dupidx)>0) warning(paste("Removing", sum(dupidx), "duplicates from data"))
m.df <- m.df0[!dupidx,]

## Only keep samples with complete, non-zero data for age and the 8 traits
completeidx = complete.cases(m.df[, c(age, bmi, traitnames)])
if(sum(!completeidx)>0) {
    warning(paste("Removing", sum(!completeidx), "records due to incomplete data"))
    for(ii in c(age, bmi, traitnames)) {
        tmpidx = complete.cases(m.df[, ii])
        if(sum(!tmpidx)>0) message(paste(ii, "has", sum(!tmpidx), "missing records"))
    }
    m.df <- m.df[completeidx, ]
}

message(paste("There are",nrow(m.df), "cases remaining in the dataset after removal of incomplete cases."))
```

***
## Flagging Outliers and Removing
```{r outliers_removal, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}

# Create variables for removal flag flags and set them == 0
trait.rem.flags = c()
for (var in traitnames) {
  names <- c(paste0(var, '_rem_flag', sep=''))
  trait.rem.flags[[names]] <- names
}
trait.rem.flags <- as.character(trait.rem.flags)
for (var in trait.rem.flags) {
  m.df[,var] <- 0
}

#g0 Removal
m.df$g0_rem_flag <- ifelse(m.df$g0 > 126, 1, 0)

# Add in any additional removals
# m.df$shbg_rem_flag <- ifelse(m.df$shbg > 200 & m.df$shbg_assay_method != 1, 1, 0)
# m.df$lh_rem_flag <- ifelse(m.df$lh > 50, 1, 0)
# m.df$fsh_rem_flag <- ifelse(m.df$fsh > 40, 1, 0)


for (var in traitnames) {
  x <- paste0(var,"_rem_flag")
  if(sum(m.df[,x]) > 0) message(paste(var, "has", sum(m.df[,x]), "records removed as outliers."))
}

m.df.removed <- filter(m.df, if_any(all_of(trait.rem.flags), ~ . == 1))
removed.path <- 'J:/endocrinology/Dunaif_Team/Dunaif_Lab/__MultiEthnic FineMap R01 HD085227/Fine Map Replication-MEGA/Analysis/MEGA GWAS/Subtyping Analysis/Control_Clustering_Removed_Outliers.txt'
#write.table(m.df.removed, removed.path, sep = '\t', row.names = F, quote = F)
message(paste0('Outliers removed from dataset written to: ', removed.path))

for (var in traitnames) {
  x <- paste0(var,"_rem_flag")
  m.df[,var] <- ifelse(m.df[,x] == 1, NA, m.df[,var])
}

## Only keep samples with complete, non-zero data for age and the 8 traits
completeidx = complete.cases(m.df[, c(age, bmi, traitnames)])
if(sum(!completeidx)>0) {
    warning(paste("Removing", sum(!completeidx), "records records containing outliers"))
    m.df <- m.df[completeidx, ]
}

message(paste("There are",nrow(m.df), "cases remaining in the dataset after outliers were removed."))


```

### Adjust BMI and 7 traits by Age, then apply reverse normal transformation
```{r adjust_normalize, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}
################################################
# Adjust BMI and 7 other traits               #
# Then apply reverse normal transformation      #
################################################

## Adjust log(BMI) for age, obtain the residual.
m.df$bmi.z <- log(m.df$bmi)
model <- lm(bmi.z ~ age, data=m.df)
m.df$bmi.z <- resid(model)

## For the other 7 traits, adjust log(trait) for age and assay method (if
## there are more than one method), obtain the residual.
for (var in traitnames) {
  method <- paste0(var, assay)
  z <- paste(var, 'z', sep=d)
  m.df[,method] <- as.factor(as.character(m.df[,method]))
  m.df[, z] <- log(m.df[, var])
  if (nlevels(m.df[,method]) > 1) {
    model <- lm(as.formula(paste(z, '~', age, '+', method)), data=m.df)
  } else {
    model <- lm(as.formula(paste(z, '~', age)), data=m.df)
  }
  m.df[, z] <-  resid(model)
}

## Apply the inverse normal transformation to the residuals
for (col in cluster_cols[3:length(cluster_cols)]) {
  #print(c(col,shapiro.test(m.df[,col])[[2]] ))
  m.df[!is.na(m.df[,col]), col] <- rntransform(m.df[!is.na(m.df[,col]), col])
}
```


```{r create_clust_df, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}
# Display number of subjects being included after removals

m.df2 <- m.df[, cluster_cols]
names(m.df2) <- c(sample, age, var_labels)
```
***

```{r correlation, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}
corrplot(cor(m.df2[3:10]), method = 'number', type = 'lower')
```

### Hierarchical Clustering
```{r perform_clustering, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}
## Hierarchical clustering on the columns
cluster_matrix <- as.matrix(na.omit(m.df2[, var_labels]))
hc <- hclust(dist(cluster_matrix, method=dist_metric), method=clust_method)

## Hierarchical clustering on the rows
df.12 <- as.matrix(na.omit(m.df2[, var_labels]))
row_hc <- hclust(dist(t(df.12), method=dist_metric), method=clust_method)

## Define clusters (k=3 means we define 3 clusters)
mycl <- cutree(hc, k=3)

## Get clustering stats
stats <- cluster.stats(dist(cluster_matrix, method=dist_metric), mycl)

## Add the cluster ID# to your data and relabel the clusters
clusters <- as.data.frame(cbind(cluster_matrix, cluster_id=mycl))

```

```{r relabel_clusters, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}
centroids <- clusters %>% 
  group_by(cluster_id) %>%
  summarise_all(funs(mean))
centroids.met <- centroids[,c('BMI', 'Ins0', 'Glu0')]
centroids.rep <- centroids[,c('SHBG', 'LH', 'FSH')]
met <- which.max(rowSums(centroids.met))
rep <- which.max(rowSums(centroids.rep))

clusters$cluster_new <- ifelse(clusters$cluster_id==met, 10,
                               ifelse(clusters$cluster_id==rep, 11, 12))
clusters$cluster_id <- ifelse(clusters$cluster_new==10, 1,
                              ifelse(clusters$cluster_new==11, 2, 3))

clusters$cluster_new = NULL  ## remove this temporary column
clusters$group <- ifelse(clusters$cluster_id==1, "Metabolic",
                         ifelse(clusters$cluster_id==2, "Reproductive", "Indeterminate"))


## Apply updated cluster ID's an updated variable 'mycl_upd' that was generated from the 
## cuttree function after pruning to 3 from the hierarchical clustering
mycl_upd <- clusters$cluster_id
```

```{r clustering_stats, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}
stats <- cluster.stats(dist(cluster_matrix, method=dist_metric), mycl_upd)

```
***

***
### HC Boxplots of Normalized Traits by Cluster
```{r boxplots, eval=TRUE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, out.width= "110%"}
################################################
# Boxplots of the traits by cluster            #
################################################

## Merge the sample_ids
clusters <- merge(clusters, m.df2, by=var_labels)
clusters.df <- clusters

hc.melt <- melt(clusters[1:9], id.var = "cluster_id")

sorted_vars <- var_labels
plot_labels <- c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG')
hc.melt$variable <- factor(hc.melt$variable, levels = c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG'))


ggplot(data = hc.melt, aes(x=variable, y=value, fill=as.factor(cluster_id))) + geom_boxplot(outlier.shape = NA, coef = 0) +
  scale_fill_manual(labels = c("Metabolic", "Reproductive", "Indeterminate"), values=c(rgb(0.81,0.3,0.25,.85), rgb(.33,.5,.77,1), 'grey'), name = "") + ylim(-2,2) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.box = "horizontal", 
        axis.title.x=element_blank()) + ylab("Z")

```


```{r updatelabel, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}
## Re-order the data frame
col_idx <- grep("cluster_id", names(clusters))
clusters <- clusters[, c(col_idx, (1:ncol(clusters))[-col_idx])]
col_idx <- grep(sample, names(clusters))
clusters <- clusters[, c(col_idx, (1:ncol(clusters))[-col_idx])]


## Write output table
#write.csv(clusters, "path/to/HC_cluster_groups.csv")

```

***

### PCA Plot of Cluster Results
```{r pca_plots, eval=TRUE, echo=FALSE, error=FALSE, message=TRUE}

################################################
# PCA plots overlay the clusters               #
################################################

### PCA PLOT ###
clusters$cluster_id <- as.factor(clusters$cluster_id)
clusters$group <- ifelse(clusters$cluster_id==1, "Metabolic", ifelse(clusters$cluster_id==2, "Reproductive", "Indeterminate"))

pca <- PCA(clusters[3:10], graph = F)

# If plot is mirrored from example, multiply PC1 by -1 to reverse image to
# maintain consistency.

#pca$ind$coord[,1] <- pca$ind$coord[,1]*-1
#pca$var$coord[,1] <- pca$var$coord[,1]*-1

fviz_pca_biplot(pca, col.ind = clusters$group, palette = c('grey', rgb(0.81,0.3,0.25,.85), rgb(.33,.5,.77,1)),
                addEllipses = TRUE, label = "var", col.var = "black", repel = TRUE,
                legend.title = "Subtype", pointsize=2,#pointshape=19,
                xlab = paste("PC1 (", round(pca$eig[1,2],2), '%)', sep=''),
                ylab = paste("PC2 (", round(pca$eig[2,2],2), '%)', sep=''),
                title = "")
#dev.off()


```

***

## Kmeans clustering

```{r Kmeans, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
m.df_km <- m.df2[3:10]
df <-scale(m.df_km)

##kmeans cluster
library(flexclust)
library(dplyr)
library(fpc)
library(DT)

set.seed(123)
clk3 <- cclust(df, k=3, method = "kmeans", dist  = "manhattan")
#barchart(clk3,legend=TRUE)

m.df_km$cluster <- clk3@cluster
m.df_km$cluster<-as.character(m.df_km$cluster)

##Label clusters by name
centroids_km <- m.df_km %>% 
  group_by(cluster) %>%
  summarise_all(funs(mean))
centroids.met_km <- centroids_km[,c('BMI', 'Ins0', 'Glu0')]
centroids.rep_km <- centroids_km[,c('SHBG', 'LH', 'FSH')]
met_km <- which.max(rowSums(centroids.met_km))
rep_km <- which.max(rowSums(centroids.rep_km))


m.df_km$cluster_new <- ifelse(m.df_km$cluster==met_km, 10,
                               ifelse(m.df_km$cluster==rep_km, 11, 12))
m.df_km$cluster_id <- ifelse(m.df_km$cluster_new==10, 1,
                              ifelse(m.df_km$cluster_new==11, 2, 3))

m.df_km$cluster_new = NULL  ## remove this temporary column
m.df_km$cluster = NULL ##remove original cluster number

m.df_km$group <- ifelse(m.df_km$cluster_id==1, "Metabolic",
                         ifelse(m.df_km$cluster_id==2, "Reproductive", "Indeterminate"))

#Merge named clusters with subject IDs 
clusters_km <- merge(m.df_km, m.df2, by=var_labels)

#write.csv(clusters_km,"path/to/Kmeans_cluster_groups.csv")


###visualize plot by PCA
res.pca <- PCA(df, graph = FALSE)

#str(res.pca)

#res.pca$ind$coord[,1] <- res.pca$ind$coord[,1]*-1
#res.pca$var$coord[,1] <- res.pca$var$coord[,1]*-1

cluster_colors_km <- c('grey', "#CF4D40D9", "#5480C4FF") 

fviz_pca_biplot(
res.pca,
axes = c(1, 2),
geom.ind = "point",
geom.var = c("arrow", "text"),
col.ind = m.df_km$group,
fill.ind = "white",
col.var = "black",
fill.var = "white",
gradient.cols = NULL,
label = "all",
invisible = "none",
repel = FALSE,
habillage = "none",
palette = cluster_colors_km,
addEllipses = TRUE,
title = "PCA - Biplot")
```

***
### Kmeans Boxplots of Normalized Traits by Cluster
```{r boxplots_km, eval=TRUE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, out.width= "110%"}
################################################
# Boxplots of the traits by cluster            #
################################################

## Merge the sample_ids
hc.melt_km <- melt(clusters_km[1:9], id.var = "cluster_id")

#sorted_vars <- var_labels
plot_labels <- c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG')
hc.melt_km$variable <- factor(hc.melt_km$variable, levels = c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG'))


ggplot(data = hc.melt_km, aes(x=variable, y=value, fill=as.factor(cluster_id))) + 
  geom_boxplot(outlier.shape = NA, coef = 0) +
  scale_fill_manual(labels = c("Metabolic", "Reproductive", "Indeterminate"), values=c(rgb(0.81,0.3,0.25,.85), rgb(.33,.5,.77,1), 'grey'), name = "") + ylim(-2,2) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.box = "horizontal", 
        axis.title.x=element_blank()) + ylab("Z")

```

## Gaussian Mixed Models clustering

```{r Gaussian, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
##gaussian mixed models cluster
library(ClusterR)

m.df_gmm <- m.df2[3:10]
data_gmm <-center_scale(m.df_gmm, mean_center = T, sd_scale = T) #center and scale data

gmm <- GMM(data_gmm, 3, dist_mode = "maha_dist", seed_mode = "random_subset", km_iter =10, em_iter = 10, verbose =F)

pr= predict_GMM(data_gmm, gmm$centroids, gmm$covariance_matrices, gmm$weights)

m.df_gmm$cluster <- pr$cluster_labels
m.df_gmm$cluster<-as.character(m.df_gmm$cluster)

##Label clusters by name
centroids_gmm <- m.df_gmm %>% 
  group_by(cluster) %>%
  summarise_all(funs(mean))
centroids.met_gmm <- centroids_gmm[,c('BMI', 'Ins0', 'Glu0')]
centroids.rep_gmm <- centroids_gmm[,c('SHBG', 'LH', 'FSH')]
met_gmm <- which.max(rowSums(centroids.met_gmm))
rep_gmm <- which.max(rowSums(centroids.rep_gmm))


m.df_gmm$cluster_new <- ifelse(m.df_gmm$cluster==met_gmm, 10,
                               ifelse(m.df_gmm$cluster==rep_gmm, 11, 12))
m.df_gmm$cluster_id <- ifelse(m.df_gmm$cluster_new==10, 1,
                              ifelse(m.df_gmm$cluster_new==11, 2, 3))

m.df_gmm$cluster_new = NULL  ## remove this temporary column
m.df_gmm$cluster = NULL ##remove original cluster number

m.df_gmm$group <- ifelse(m.df_gmm$cluster_id==1, "Metabolic",
                         ifelse(m.df_gmm$cluster_id==2, "Reproductive", "Indeterminate"))

#Merge named clusters with subject IDs 
clusters_gmm <- merge(m.df_gmm, m.df2, by=var_labels)

#write.csv(clusters_gmm,"path/to/GMM_cluster_groups.csv")

cluster_colors_gmm <- c('grey', "#CF4D40D9", "#5480C4FF") 

fviz_pca_biplot(
res.pca,
axes = c(1, 2),
geom.ind = "point",
geom.var = c("arrow", "text"),
col.ind = m.df_gmm$group,
fill.ind = "white",
col.var = "black",
fill.var = "white",
gradient.cols = NULL,
label = "all",
invisible = "none",
repel = FALSE,
habillage = "none",
palette = cluster_colors_gmm,
addEllipses = TRUE,
title = "PCA - Biplot")
```

***
### Gaussain Boxplots of Normalized Traits by Cluster

```{r boxplots_gmm, eval=TRUE, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, out.width= "110%"}
################################################
# Boxplots of the traits by cluster            #
################################################

## Merge the sample_ids
hc.melt_gmm <- melt(clusters_gmm[1:9], id.var = "cluster_id")

#sorted_vars <- var_labels
plot_labels <- c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG')
hc.melt_gmm$variable <- factor(hc.melt_gmm$variable, levels = c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG'))


ggplot(data = hc.melt_gmm, aes(x=variable, y=value, fill=as.factor(cluster_id))) + 
  geom_boxplot(outlier.shape = NA, coef = 0) +
  scale_fill_manual(labels = c("Metabolic", "Reproductive", "Indeterminate"), values=c(rgb(0.81,0.3,0.25,.85), rgb(.33,.5,.77,1), 'grey'), name = "") + ylim(-2,2) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.box = "horizontal", 
        axis.title.x=element_blank()) + ylab("Z")

```

## Compare 3 method of clustering

```{r compare cluster groups, echo=FALSE}

clusters_compare =  merge(clusters[2:10], clusters_km[1:9], by=var_labels) %>% 
  merge(clusters_gmm, by= var_labels)  %>% rename(clusters_hc = cluster_id.x, 
                                                  clusters_km = cluster_id.y, 
                                                  clusters_gmm = cluster_id)
clusters_compare$group = NULL

consensus = function(x, len=3) {
    temp = sort(table(x), decreasing=T)
    ifelse(temp[1]>len/2, names(temp)[1], NA)
}

clusters_consensus= cbind(clusters_compare, consensus=apply(clusters_compare[9:11], 1, consensus))

```

# Cross Tables- Consensus Result vs each method

```{r Consensus Crosstables, echo=FALSE, out.width="50%"}

clusters_consensus$clusters_hc <- ifelse(clusters_consensus$clusters_hc==1, "1-Met",
                                         ifelse(clusters_consensus$clusters_hc==2, "2-Rep", 
                                                ifelse(clusters_consensus$clusters_hc==3, "3-Ind", "NA")))
clusters_consensus$clusters_km <- ifelse(clusters_consensus$clusters_km==1, "1-Met",
                                         ifelse(clusters_consensus$clusters_km==2, "2-Rep", 
                                                ifelse(clusters_consensus$clusters_km==3, "3-Ind", "NA")))
clusters_consensus$clusters_gmm <- ifelse(clusters_consensus$clusters_gmm==1, "1-Met",
                                          ifelse(clusters_consensus$clusters_gmm==2, "2-Rep",
                                                 ifelse(clusters_consensus$clusters_gmm==3, "3-Ind", "NA")))
clusters_consensus$consensus <- ifelse(clusters_consensus$consensus==1, "1-Met",
                                       ifelse(clusters_consensus$consensus==2, "2-Rep",
                                              ifelse(clusters_consensus$consensus==3, "3-Ind", "NA")))

gmodels::CrossTable(clusters_consensus$consensus, clusters_consensus$clusters_hc, digits=2, prop.r = TRUE, prop.c =TRUE,
           prop.t= FALSE, prop.chisq = FALSE, chisq =FALSE, fisher =FALSE, missing.include=FALSE)

gmodels::CrossTable(clusters_consensus$consensus, clusters_consensus$clusters_km, digits=2, prop.r = TRUE, prop.c =TRUE,
           prop.t= FALSE, prop.chisq = FALSE, chisq =FALSE, fisher =FALSE, missing.include=FALSE)

gmodels::CrossTable(clusters_consensus$consensus, clusters_consensus$clusters_gmm, digits=2, prop.r = TRUE, prop.c =TRUE,
           prop.t= FALSE, prop.chisq = FALSE, chisq =FALSE, fisher =FALSE, missing.include=FALSE)

cluster_colors = c('#CF4D40D9','#5480C4FF','grey')
```
## Visualize agreement of each method vs determined Consensus

```{r compare plots, out.width="50%"}
ggplot(clusters_consensus, aes(fill = clusters_hc, x=consensus)) +
  geom_bar() +
  scale_fill_manual(values = cluster_colors) +
   ggtitle("Consensus vs HC Clustering")

ggplot(clusters_consensus, aes(fill = clusters_km, x=consensus)) +
  geom_bar() +
  scale_fill_manual(values = cluster_colors) +
   ggtitle("Consensus vs Kmeans Clustering")

ggplot(clusters_consensus, aes(fill = clusters_gmm, x=consensus)) +
  geom_bar() +
  scale_fill_manual(values = cluster_colors) +
  ggtitle("Consensus vs GMM Clustering")

```

## Plot Traits of Consensus subjects and NA (non-Consensus)

```{r boxplots consensus, echo=FALSE}

con = clusters_consensus[c(1:8, 14)]

con.melt <- melt(con, id.var = "consensus")

#sorted_vars <- var_labels
#plot_labels <- c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG')
con.melt$variable <- factor(con.melt$variable, levels = c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG'))


ggplot(data = con.melt, aes(x=variable, y=value, fill=as.factor(consensus))) + geom_boxplot(outlier.shape = NA, coef = 0) +
  scale_fill_manual(labels = c("Metabolic", "Reproductive", "Indeterminate"), values=c(rgb(0.81,0.3,0.25,.85), rgb(.33,.5,.77,1), 'grey'), name = "") + ylim(-3.4,3.4) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.box = "horizontal", 
        axis.title.x=element_blank()) + ylab("Z") +
  ggtitle("Traits from Clustering Consensus")
```

## Visualize Traits of HC subjects that did form a consensus (Match) vs subjects that did not form a consensus (Non-Match)

```{r non-consensus HC, echo=FALSE}
HC_nonCon = clusters_consensus %>% mutate(
  hc_non= ifelse(clusters_hc=="1-Met" & consensus!="1-Met", 1,
                 ifelse(clusters_hc=="2-Rep" & consensus!="2-Rep", 1,
                        ifelse(clusters_hc=="3-Ind" & consensus!="3-Ind", 1, 0)))
)

HC_nonCon = HC_nonCon[c(12,9,14,15)]

HC_nonCon =  merge(HC_nonCon, m.df[c(1,3,4,6,8,10,12,14,16)], by="sample_id")

HC_nonCon = rename(HC_nonCon, BMI=bmi, Glu0=g0, Ins0=i0, DHEAS=dheas, SHBG=shbg, LH=lh, FSH=fsh)
HC_nonCon$hc_non = ifelse(HC_nonCon$hc_non ==1, "Non-match", 
                          ifelse(HC_nonCon$hc_non ==0, "Match", NA))
```

```{r plot non-hc_consensus, echo=FALSE}
col= c('#CF4D40D9', '#5480C4FF', 'grey', '#CF4D40D9', '#5480C4FF', 'grey')

ggplot(data = subset(HC_nonCon, !is.na(hc_non)), aes(y = BMI, x = clusters_hc, fill = clusters_hc)) + 
  geom_boxplot(fill=col) +
  ggtitle("HC and Consensus match vs non-match") +
  facet_wrap("hc_non") +
  xlab("") 

ggplot(data = subset(HC_nonCon, !is.na(hc_non)), aes(y = T, x = clusters_hc, fill = clusters_hc)) + 
  geom_boxplot(fill=col) +
  ggtitle("HC and Consensus match vs non-match") +
  facet_wrap("hc_non") +
  xlab("") 

ggplot(data = subset(HC_nonCon, !is.na(hc_non)), aes(y = Ins0, x = clusters_hc, fill = clusters_hc)) + 
  geom_boxplot(fill=col) +
  ggtitle("HC and Consensus match vs non-match") +
  facet_wrap("hc_non") +
  ylim(0,130) +
  xlab("") 

ggplot(data = subset(HC_nonCon, !is.na(hc_non)), aes(y = Glu0, x = clusters_hc, fill = clusters_hc)) + 
  geom_boxplot(fill=col) +
  ggtitle("HC and Consensus match vs non-match") +
  facet_wrap("hc_non") +
  xlab("") 

ggplot(data = subset(HC_nonCon, !is.na(hc_non)), aes(y = DHEAS, x = clusters_hc, fill = clusters_hc)) + 
  geom_boxplot(fill=col) +
  ggtitle("HC and Consensus match vs non-match") +
  facet_wrap("hc_non") +
  xlab("") 

ggplot(data = subset(HC_nonCon, !is.na(hc_non)), aes(y = SHBG, x = clusters_hc, fill = clusters_hc)) + 
  geom_boxplot(fill=col) +
  ggtitle("HC and Consensus match vs non-match") +
  facet_wrap("hc_non") +
  xlab("") 

ggplot(data = subset(HC_nonCon, !is.na(hc_non)), aes(y = LH, x = clusters_hc, fill = clusters_hc)) + 
  geom_boxplot(fill=col) +
  ggtitle("HC and Consensus match vs non-match") +
  facet_wrap("hc_non") +
  xlab("") 

ggplot(data = subset(HC_nonCon, !is.na(hc_non)), aes(y = FSH, x = clusters_hc, fill = clusters_hc)) + 
  geom_boxplot(fill=col) +
  ggtitle("HC and Consensus match vs non-match") +
  facet_wrap("hc_non") +
  xlab("") 
```

```{r strong consensus, echo=FALSE}
str_consensus= clusters_consensus %>% mutate(str_consensus=
  ifelse(clusters_hc=="1-Met" & clusters_km=="1-Met" & clusters_gmm=="1-Met", "1-Met",
         ifelse(clusters_hc=="2-Rep" & clusters_km=="2-Rep" & clusters_gmm=="2-Rep", "2-Rep",
                ifelse(clusters_hc=="3-Ind" & clusters_km=="3-Ind" & clusters_gmm=="3-Ind", "3-Ind", NA))))

```

```{r str_consensus Crosstables, echo=FALSE, out.width="50%"}

gmodels::CrossTable(str_consensus$str_consensus, str_consensus$clusters_hc, digits=2, prop.r = TRUE, prop.c =TRUE,
           prop.t= FALSE, prop.chisq = FALSE, chisq =FALSE, fisher =FALSE, missing.include=FALSE)

gmodels::CrossTable(str_consensus$str_consensus, str_consensus$clusters_km, digits=2, prop.r = TRUE, prop.c =TRUE,
           prop.t= FALSE, prop.chisq = FALSE, chisq =FALSE, fisher =FALSE, missing.include=FALSE)

gmodels::CrossTable(str_consensus$str_consensus, str_consensus$clusters_gmm, digits=2, prop.r = TRUE, prop.c =TRUE,
           prop.t= FALSE, prop.chisq = FALSE, chisq =FALSE, fisher =FALSE, missing.include=FALSE)
```

```{r compare plots str, out.width="50%"}
ggplot(str_consensus, aes(fill = clusters_hc, x=str_consensus)) +
  geom_bar() +
  scale_fill_manual(values = cluster_colors) +
   ggtitle("Strong Consensus vs HC Clustering")

ggplot(str_consensus, aes(fill = clusters_km, x=str_consensus)) +
  geom_bar() +
  scale_fill_manual(values = cluster_colors) +
   ggtitle("Strong Consensus vs Kmeans Clustering")

ggplot(str_consensus, aes(fill = clusters_gmm, x=str_consensus)) +
  geom_bar() +
  scale_fill_manual(values = cluster_colors) +
  ggtitle("Strong Consensus vs GMM Clustering")

```

```{r boxplot strong consensus, echo=FALSE}

str_con = str_consensus[c(1:8, 15)]

str_con.melt <- melt(str_con, id.var = "str_consensus")

str_con.melt$variable <- factor(str_con.melt$variable, levels = c('Ins0', 'BMI', 'Glu0', 'DHEAS', 'T', 'FSH', 'LH', 'SHBG'))


ggplot(data = str_con.melt, aes(x=variable, y=value, fill=as.factor(str_consensus))) + geom_boxplot(outlier.shape = NA, coef = 0) +
  scale_fill_manual(labels = c("Metabolic", "Reproductive", "Indeterminate"), values=c(rgb(0.81,0.3,0.25,.85), rgb(.33,.5,.77,1), 'grey'), name = "") + ylim(-3.4,3.4) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.box = "horizontal", 
        axis.title.x=element_blank()) + ylab("Z") +
  ggtitle("Traits from Strong Consensus")

write.csv(str_consensus, "path/to/Consensus_clusters.csv")
```

